---
title: "Grammar of graphics"
subtitle: "Lecture 3"
date: "September 6, 2022"
format: revealjs
---

# Warm up

```{r}
#| echo: false

library(countdown)
```

## Announcements

-   Office hours start this week
-   "Graded" AEs also start this week -- commit and push to your repo within 3 days of AE

## **Examining data visualization** {.smaller}

::: columns
::: {.column width="50%"}
Discuss the following for the visualization.

-   What is the visualization trying to show?

-   What is effective, i.e. what is done well?

-   What is ineffective, i.e. what could be improved?

-   What are you curious about after looking at the visualization?

```{r}
#| echo: false

countdown(minutes = 4, left = 0, color_text = "#325b74")
```
:::

::: {.column width="50%"}
![Source: [Twitter](https://twitter.com/reina_sabah/status/1291509085855260672)](images/03/female-heights.png){fig-align="center"}
:::
:::

## Questions from the prepare materials?

```{=html}
<iframe src="https://app.sli.do/event/wCtRaWXjUCSQGm1untvmE5/embed/polls/8a814dd9-a20e-4a10-911c-cde825e98c35" width="900" height="600"></iframe>
```
# Application exercise

## `ae-01`

::: appex
-   Go to the course [GitHub org](https://github.com/sta199-f22-1) and find your `ae-01` (repo name will be suffixed with your GitHub name).
-   Clone the repo in your container, open the Quarto document in the repo, and follow along and complete the exercises.
-   Render, commit, and push your edits by the AE deadline -- 3 days from today.
:::

## Recap of AE {.smaller}

-   Construct plots with `ggplot()`.
-   Layers of ggplots are separated by `+`s.
-   The formula is (almost) always as follows:

```{r}
#| eval: false

ggplot(DATA, aes(x = X-VAR, y = Y-VAR, ...)) +
  geom_XXX()
```

-   Aesthetic attributes of a geometries (color, size, transparency, etc.) can be **map**ped to variables in the data or **set** by the user, e.g. `color = binary` vs. `color = "pink"`.
-   Use `facet_wrap()` when faceting (creating small multiples) by one variable and `facet_grid()` when faceting by two variables.
