---
title: "Visualizing various types of data"
subtitle: "Lecture 4"
date: "September 8, 2022"
format: revealjs
---

# Warm up

```{r}
#| echo: false

library(countdown)
```

## **Identifying variable types** {.smaller}

::: question
Identify the type of each of the following variables.

-   Favorite food
-   Number of classes you're taking this semester
-   Zip code
:::

## Questions from the prepare materials?

```{=html}
<iframe src="https://app.sli.do/event/9mqwqdqbgesgLASUHGpwU8/embed/polls/52941ba5-0668-4cad-9b86-82505db6b862" width="900" height="600"></iframe>
```
## **The way data is displayed matters**

::: question
What do these three plots show?
:::

![Source: [#barbarplots](https://cogtales.wordpress.com/2016/06/06/congratulations-barbarplots/)](images/04/bar-plot-meme.png){fig-alt="Three representations of the same data: Bar plot of means, box plots, histograms." fig-align="center"}

# Application exercise

## `ae-02-bechdel-dataviz`

::: appex
-   Go to the course [GitHub org](https://github.com/sta199-f22-1) and find your `ae-01-bechdel-dataviz` (repo name will be suffixed with your GitHub name).
-   Clone the repo in your container, open the Quarto document in the repo, and follow along and complete the exercises.
-   Render, commit, and push your edits by the AE deadline -- 3 days from today.
:::

## Recap of AE {.smaller}

-   Construct plots with `ggplot()`.
-   Layers of ggplots are separated by `+`s.
-   The formula is (almost) always as follows:

```{r}
#| eval: false

ggplot(DATA, aes(x = X-VAR, y = Y-VAR, ...)) +
  geom_XXX()
```

-   Aesthetic attributes of a geometries (color, size, transparency, etc.) can be **map**ped to variables in the data or **set** by the user, e.g. `color = binary` vs. `color = "pink"`.
-   Use `facet_wrap()` when faceting (creating small multiples) by one variable and `facet_grid()` when faceting by two variables.
