---
title: "Grammar of data wrangling"
subtitle: "Lecture 5"
date: "September 13, 2022"
format: revealjs
---

# Warm up

```{r}
#| echo: false

library(countdown)
```

## While you wait for class to begin...

::: nonincremental
-   Clone your `ae-03` repo.
-   Any questions from prepare materials? You can also upvote others' questions.
:::

```{=html}
<iframe src="https://app.sli.do/event/7mtcYfUfiJavFcqsPg2jWN" height="900" width="600" frameBorder="0" style="min-height: 560px;" title="Slido"></iframe>
```
## Announcements

-   Request videos for missed classes via the **video request form**
-   Ask course questions **on Slack**
    -   Using code formatting and proper indentation

    -   Taking screenshots

    -   Checking for previous questions

## Coding style + workflow

::: incremental
-   Avoid long lines of code.

    -   We should be able to see [all]{.underline} of your code in the PDF document you submit.

-   Label code chunks.

    -   Do **not** put spaces in the code-chunk `label`s.

-   Render, commit, and push regularly.

    -   Think about it like clicking to save regularly as you type a report.
:::

# Application exercise

## `ae-03`

::: appex
-   Go to the course [GitHub org](https://github.com/sta199-f22-1) and find your `ae-03` (repo name will be suffixed with your GitHub name).
-   Clone the repo in your container, open the Quarto document in the repo, and follow along and complete the exercises.
-   Render, commit, and push your edits by the AE deadline (3 days from today).
:::

## Recap of AE {.smaller}

-   The pipe operator, `|>`, can be read as "and then".
-   The pipe operator passes what comes before it into the function that comes after it as the first argument in that function.

```{r}
sum(1, 2)
1 |> 
  sum(2)
```

-   Always use a line break after the pipe, and indent the next line of code.
    -   Just like always use a line break between layers of ggplots, after `+`, and indent the next line.
