---
title: "Inference overview"
subtitle: "Lecture 22"
date: "November 15, 2022"
format: revealjs
---

# Warm up

```{r}
#| echo: false

library(countdown)
```

## While you wait for class to begin...

::: callout-important
Come up with at least one question you'd like to see answered in the Exam 1 review on Thursday.

Go to [slido.com](https://www.slido.com/) and submit your question with the code **#sta199exam2**.

You can also upvote others' questions.
:::

## Announcements {.smaller}

-   My office hours today 3:30-5pm on Zoom: [bit.ly/minezoom](https://bit.ly/minezoom).

-   Lab 6 due Friday night (11:59 pm) -- but I strongly recommend you finish and submit it before the exam begins!

-   Exam 2 is released on Thursday at noon and is due at 2pm on Monday.

    -   No TA OH during the exam.

    -   I will have OH 4-5pm on Friday (on Zoom).

    -   Any clarification questions must be **emailed to me** only.

    -   No Slack use during the exam, even about non-exam related questions.

## Exam study tips

-   Take notes that you can reference
-   Review feedback given to you on graded work
-   Revisit AEs
-   Review suggested solutions
-   Ask questions

## Topics for review

::: question
Come up with at least one question you'd like to see answered in the Exam 2 review on Thursday.

Go to [slido.com](https://www.slido.com/) and submit your question with the code **#sta199exam2**.

You can also upvote others' questions.
:::

```{r}
#| echo: false

countdown(minutes = 3)
```

# Application exercise

## Recap of inference ideas {.smaller}

-   A hypothesis test is a statistical technique used to evaluate competing claims (null and alternative hypotheses) using data.
-   We simulate a null distribution using our original data.
-   We use our sample statistic and direction of the alternative hypothesis to calculate the p-value.
-   We use the p-value to determine conclusions about the alternative hypotheses.

## `ae-19`

::: appex
-   Go to the course [GitHub org](https://github.com/sta199-f22-1) and find your `ae-19` (repo name will be suffixed with your GitHub name).
-   Clone the repo in your container, open the Quarto document in the repo, and follow along and complete the exercises.
-   Render, commit, and push your edits by the AE deadline -- 3 days from today.
:::
