---
title: "Exam 2 review"
subtitle: "Lecture 23"
date: "November 17, 2022"
format: revealjs
---

# Warm up

## Announcements {.smaller}

-   Exam 2 is released on today at noon and is due at 2pm on Monday.

    -   No TA OH during the exam.

    -   I will have OH 4-5pm on Friday (on Zoom).

    -   Any clarification questions must be **emailed to me** only.

    -   No Slack use during the exam, even about non-exam related questions.

-   Answer keys:

    -   HW 4 and HW 5 feedback: Keys both posted, feedback coming soon.

    -   All lab keys also posted.

    -   AE 16 key missing complete answers, will post after class.

# Review questions

## Questions: Grouping {.smaller}

::: question
Does order of variables in `group_by()` matter?
:::

```{r}
#| message: false

library(palmerpenguins)
library(tidyverse)
```

::: columns
::: {.column width="50%"}
```{r}
penguins |>
  group_by(species, sex) |>
  summarize(mean_bm = mean(body_mass_g))
```
:::

::: {.column width="50%"}
```{r}
penguins |>
  group_by(sex, species) |>
  summarize(mean_bm = mean(body_mass_g))
```
:::
:::

## Questions: Factors

::: question
-   When will we use factors and how does that make a difference in the data?

-   When do you use `fct_relevel()` versus `fct_reorder()`?

-   How to use `case_when()` function and the proper use of forcats functions like `fct_relevel()` , `fct_reorder()`, `fct_other()`?
:::

Review: <https://r4ds.hadley.nz/factors.html>.

## Questions: LaTeX / equations

::: question
-   What is the name of the math symbol text we use to write equations? Is there a cheat sheet with the shortcuts for each symbol?
-   Do we have to use LaTeX for our equations?
:::

::: columns
::: {.column width="50%"}
$H_0:\mu_1 - \mu_2 = 0$

$H_A: \mu_1 - \mu_2 \ne 0$

    $H_0:\mu_1 - \mu_2 = 0$

    $H_A: \mu_1 - \mu_2 \ne 0$
:::

::: {.column width="50%"}
H0: mu1 - mu2 = 0

HA: mu1 - mu2 ≠ 0
:::
:::

## Questions: Ethics

::: question
Would love to review data ethics and how to answer questions about ethical issues with any dataset.
:::

Review: The videos from the Ethics module.

## Questions: Miscellaneous

::: question
What does geom_smooth(method = "loess") do?
:::

. . .

Fits a non-linear model to the data, a smooth curve.

. . .

::: question
How do we know how much to round by?
:::

. . .

Round as much as it makes sense in the context of the data.
Avoid rounding in interim steps.

## Question: Inference

::: question
How do we decide whether to use bootstrap, simulate, or permute in the `generate()` step of inference?
:::

-   Bootstrap: For constructing bootstrap intervals or for testing for a single mean ($H0: \mu_0 = 5$)

-   Simulate: For testing for a single proportion ($H_0: p_0 = 0.3$)

-   Permute: For testing for independence, i.e., for testing for differences in means or proportions across groups (or whether one is less/greater than the other)

# Application exercise

## `ae-19`

::: appex
-   Go to the course [GitHub org](https://github.com/sta199-f22-1) and find your `ae-19` (repo name will be suffixed with your GitHub name).
-   Clone the repo in your container, open the Quarto document in the repo, and follow along and complete the exercises.
-   You should have already pushed updates on Tuesday, so you should be good for submission.
:::
