---
title: "Quantifying uncertainty"
categories: 
  - Application exercise
editor: visual
editor_options: 
  chunk_output_type: console
---

An article in the The Asheville Citizen-Times published in the summer of 2020 claims that the average price per guest (ppg) for properties in Asheville is \$60 on Airbnb.
To evaluate their claim we will use a dataset on 50 randomly selected Asheville Airbnb listings in June 2020.
These data can be found in `data/asheville.csv`.

Let's load the packages we'll use first.

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(tidymodels)
```

And then the data.

```{r}
#| label: load-data
#| message: false

# add code here
```

# Hypotheses

-   **Your turn:** Write out the correct null and alternative hypothesis. Do this in both words *and* in proper notation.

*Add response here.*

# Observed data

Our goal is to use calculate the probability of a sample statistic at least as extreme as the one observed in our data if in fact the null hypothesis is true.

-   **Demo:** Calculate and report the sample statistic below using proper notation.

```{r}
#| label: point-estimate

# add code here
```

$$\bar{x} = 76.6$$

# The null distribution

Let's use simulation-based methods to conduct the hypothesis test specified above.

## Generate

We'll start by generating the null distribution.

-   **Demo:** Generate the null distribution?

```{r}
#| label: set-seed

set.seed(4321)
```

```{r}
#| label: null-dist-generate

# add code here
```

-   **Your turn:** Take a look at `null_dist`. What does each element in this distribution represent?

```{r}
#| label: null-dist-view

# add code here
```

## Visualize

-   **Question:** Before you visualize the distribution of `null_dist` -- at what value would you expect this distribution to be centered? Why?

*Add response here.*

-   **Demo:** Create an appropriate visualization for your null distribution. Does the center of the distribution match what you guessed in the previous question?

```{r}
#| label: null-dist-visualize

# add code here
```

-   **Demo:** Now, add a vertical red line on your null distribution that represents your sample statistic.

```{r}
#| label: null-dist-annotate

# add code here
```

-   **Question:** Based on the position of this line, does your observed sample mean appear to be an unusual observation under the assumption of the null hypothesis?

*Add response here.*

# p-value

Above, we eyeballed how likely/unlikely our observed mean is.
Now, let's actually quantify it using a p-value.

-   **Question:** What is a p-value?

*Add response here.*

-   **Demo:** Visualize the p-value.

```{r}
#| label: p-value-visualize

# add code here
```

-   **Your turn:** What is the p-value?

*Add response here.*

# Conclusion

-   **Your turn:** What is the conclusion of the hypothesis test based on the p-value you calculated? Make sure to frame it in context of the data and the research question. Use a significance level of 5% to make your conclusion.

*Add response here.*

-   **Demo:** Interpret the p-value in context of the data and the research question.

*Add response here.*

# Get real...

-   **Question:** What we did above was a "toy example" to illustrate hypothesis test. What would you change to make this a real, more robust analysis?

*Add response here.*

-   **Demo:** Work through the analysis again with these changes.

```{r}
#| label: hypothesis-test

# add-code-here
```
