---
title: "Yankee Candle reviews on Amazon"
categories: 
  - Application exercise
editor: visual
---

## Part 1 - Data scraping

Do this part in `yankee-candle-scrape.R`.

## Part 2 - Data analysis

Let's start by loading the packages we will need:

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(lubridate)
```

-   Load the data you saved into the `data` folder and name it `yc_reviews_raw`.

```{r}
#| label: load-data
#| message: false

# add code here
```

-   **Your turn (5 minutes):** Our goal is to plot the number of reviews mentioning "no scent" or "no smell" per week, limiting our plot to reviews posted in 2022. Take a look at the data you loaded and outline how you can transform it to prepare it for such a plot. Your response should be bullet points that can roughly be translated to steps in a data wrangling pipeline.

*Add your response here.*

-   **Demo:** Create the plot described above.

```{r}
#| label: plot-reviews

# add code here
```

-   **Your turn (3 minutes):** Outline how this exercise could be expanded to make a plot like [this one](https://twitter.com/nick_beauchamp/status/1473509106187223043) that compares number of reviews mentioning no scent/smell and number of COVID cases.

*Add your response here.*
