---
title: "Modelling loan interest rates"
subtitle: "Suggested answers"
categories: 
  - Application exercise
  - Answers
editor: visual
---

In this application exercise we will be studying loan interest rates.
The dataset is one you've come across before in your reading -- the dataset about loans from the peer-to-peer lender, Lending Club, from the **openintro** package.
We will use **tidyverse** and **tidymodels** for data exploration and modeling, respectively.

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(tidymodels)
library(openintro)
```

Before we use the dataset, we'll make a few transformations to it.

```{r}
loans <- loans_full_schema %>%
  mutate(
    credit_util = total_credit_utilized / total_credit_limit,
    bankruptcy  = as.factor(if_else(public_record_bankrupt == 0, 0, 1)),
    verified_income = droplevels(verified_income),
    homeownership = str_to_title(homeownership),
    homeownership = fct_relevel(homeownership, "Rent", "Mortgage", "Own")
    ) %>%
  rename(credit_checks = inquiries_last_12m) %>%
  select(interest_rate, verified_income, debt_to_income, credit_util, bankruptcy, term, credit_checks, issue_month, homeownership) 
```

Here is a glimpse at the data:

```{r}
#| label: glimpse-penguins

glimpse(penguins)
```

# Interest rate vs. credit utilization ratio

The regression model for interest rate vs. credit utilization is as follows.

```{r}
#| label: fit-rate-util

rate_util_fit <- linear_reg() |>
  fit(interest_rate ~ credit_util, data = loans)

tidy(rate_util_fit)
```

And here is the model visualized:

```{r}
ggplot(loans, aes(x = credit_util, y = interest_rate)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm")
```

-   **Your turn:** What is the estimated interest rate for a loan applicant with credit utilization of 0.8, i.e. someone whose total credit balance is 80% of their total available credit?

```{r}
#| label: predict-80

# add code here
```

# Interest rate vs. homeownership

Next we predict interest rates from homeownership, which is a categorical predictor with three levels:

```{r}
levels(loans$homeownership)
```

-   **Demo:** Fit the linear regression model to predict interest rate from homeownership and display a tidy summary of the model. Write the estimated model output below.

```{r}
#| label: fit-rate-home

# add code here
```

-   **Your turn:** Interpret each coefficient in context of the problem.

    -   Intercept: *Add response here.*

    -   Slopes: *Add response here.*

# Interest rate vs. credit utilization and homeownership

## Main effects model

-   **Demo:** Fit a model to predict interest rate from credit utilization and homeownership, **without** an interaction effect between the two predictors. Display the summary output and write out the estimated regression equation.

```{r}
#| label: rate-util-home-fit

# add code here
```

$$
add~math~text~here
$$

-   **Demo:** Write the estimated regression equation for loan applications from each of the homeownership groups separately.
    -   Rent: $add~math~text~here$
    -   Mortgage: $add~math~text~here$
    -   Own: $add~math~text~here$
-   **Question:** How does the model predict the interest rate to vary as credit utilization varies for loan applicants with different homeownership status. Are the rates the same or different?

*Add response here.*

## Interaction effects model

-   **Demo:** Fit a model to predict interest rate from credit utilization and homeownership, **with** an interaction effect between the two predictors. Display the summary output and write out the estimated regression equation.

```{r}
#| label: rate-util-home-int-fit

# add code here
```

$$
add~math~text~here
$$

-   **Demo:** Write the estimated regression equation for loan applications from each of the homeownership groups separately.
    -   Rent: $add~math~text~here$
    -   Mortgage: $add~math~text~here$
    -   Own: $add~math~text~here$
-   **Question:** How does the model predict the interest rate to vary as credit utilization varies for loan applicants with different homeownership status. Are the rates the same or different?

*Add response here.*

## Choosing a model

Rule of thumb: **Occam's Razor** - Don't overcomplicate the situation!
We prefer the *simplest* best model.

```{r}
#| label: glance

# add code here
```

-   **Review:** What is R-squared? What is adjusted R-squared?

*Add response here.*

-   **Question:** Based on the adjusted $R^2$s of these two models, which one do we prefer?

*Add response here.*

# Another model to consider

-   **Your turn:** Let's add one more model to the variable -- issue month. Should we add this variable to the interaction effects model from earlier?

```{r}
# add code here
```

*Add response here.*
