---
title: "Yankee Candle reviews on Amazon"
subtitle: "Suggested answers"
categories: 
  - Application exercise
  - Answers
editor: visual
---

## Part 1 - Data scraping

See [`yankee-candle-scrape-A.R`](https://github.com/sta199-f22-1/sta199-f22-1.github.io/blob/main/ae/_yankee-candle-scrape-A.R) for suggested scraping code.

## Part 2 - Data analysis

Let's start by loading the packages we will need:

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(lubridate)
```

-   Load the data you saved into the `data` folder and name it `yc_reviews_all`.

```{r}
#| label: load-data
#| message: false

yc_reviews_all <- read_csv("data/yc-reviews-all.csv")
```

-   **Your turn (5 minutes):** Our goal is to plot the number of reviews or titles mentioning "no scent" or "no smell" per week. Take a look at the data you loaded and outline how you can transform it to prepare it for such a plot. Your response should be bullet points that can roughly be translated to steps in a data wrangling pipeline.
    -   Extract date of review.
    -   Figure out which week of the year the date corresponds to.
    -   Look for "no scent" or "no smell" in the titles or reviews.
    -   Count how many titles or reviews have these phrases mentioned for each week.
    -   Plot the number of mentions by week against week number.
-   **Demo:** Create the plot described above.

```{r}
#| label: plot-reviews

yc_reviews_all |>
  mutate(
    date = mdy(country_date),
    week = week(date),
    review = str_to_lower(review),
    title = str_to_lower(title),
    no_scent = case_when(
      str_detect(review, "no scent") | str_detect(title, "no scent") ~ "mentioned",
      str_detect(review, "no smell") | str_detect(title, "no smell") ~ "mentioned",
      TRUE ~ "not mentioned"
    )
  ) |>
  count(week, no_scent) |>
  filter(no_scent == "mentioned") |>
  ggplot(aes(x = week, y = n, group = 1)) +
  geom_line()
```

-   **Your turn (3 minutes):** Outline how this exercise could be expanded to make a plot like [this one](https://twitter.com/nick_beauchamp/status/1473509106187223043) that compares number of reviews mentioning no scent/smell and number of COVID cases.

Scrape data going further back and join it with weekly COVID cases data from another source and plot both.
